---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

#### Load library

```{r}
library(tidyverse)
```

## Loading and preprocessing the data

```{r}
if(!dir.exists('data')) {
    dir.create('data')
    unzip('activity.zip', exdir = 'data')
}
data <- read_csv('data/activity.csv')
data
```
## What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
```{r}
steps_per_day <- data %>%
    group_by(date) %>%
    summarize(steps = sum(steps, na.rm = TRUE))
steps_per_day
```
Make histogram of the total number of steps taken each day.
```{r}
steps_per_day %>% ggplot(aes(steps)) +
    geom_histogram(bins = 6, fill = '#f7e599', color = '#f2a600') +
    labs(title = 'Histogram of the total number of steps taken each day',
         x = 'Steps',
         y = 'Count') +
    theme_bw()
```

Mean and median number of steps
```{r}
mean(steps_per_day$steps)
median(steps_per_day$steps)
```

## What is the average daily activity pattern?
Calculate average number of steps for each interval
```{r}
mean_steps_interval <- data %>%
    group_by(interval) %>%
    summarize(steps = mean(steps, na.rm = TRUE))
```
Make a time series plot of the 5 minutes interval and the avegare of steps taken.
```{r}
mean_steps_interval %>%
    ggplot(aes(interval, steps)) +
    geom_line(lwd = 1, color = '#00c9a1') +
    labs(title = 'Average steps of each interval',
         x = 'Interval',
         y = 'Steps') +
    theme_bw()
```

Interval contains the maximum number of steps
```{r}
m <- which.max(mean_steps_interval$steps)
mean_steps_interval[m,]
```

## Imputing missing values
Calculate total number of missing values
```{r}
sum(is.na(data$steps))
```
Replace missing values with the mean for 5-minutes interval
```{r}
new_data <- data %>%
    group_by(interval) %>%
    mutate(steps = if_else(is.na(steps), mean(steps, na.rm = TRUE), steps)) %>%
    ungroup()
```
Calculate the total number of steps taken per day with `new_data`
```{r}
new_steps_per_day <- new_data %>%
    group_by(date) %>%
    summarize(steps = sum(steps, na.rm = TRUE))
new_steps_per_day
```
Make histogram of the total number of steps taken each day with `new_data`
```{r}
new_steps_per_day %>% ggplot(aes(steps)) +
    geom_histogram(bins = 6, fill = '#f7e599', color = '#f2a600') +
    labs(title = 'Histogram of the total number of steps taken each day',
         x = 'Steps',
         y = 'Count') +
    theme_bw()
```

Mean and median number of steps with `new_data`
```{r}
summary(new_steps_per_day$steps)
```
Imputing missing data increase mean and median number of steps per day
```{r}
summary(steps_per_day$steps)
```
## Are there differences in activity patterns between weekdays and weekends?
Create `weekd` variables in dataset with two levels 'weekday' and 'weekend'
```{r}
new_data <- new_data %>%
    mutate(weekd = factor(
        if_else(weekdays(date, abbreviate=TRUE) %in% c('Sat','Sun'), 'weekend', 'weekday')
    ))
```
Calculate the average number of steps for each 5-minute interval on weekdays or weekends, make time series plot.
```{r}
new_data %>%
    group_by(interval, weekd) %>%
    summarize(steps = mean(steps)) %>%
    ggplot(aes(interval, steps)) +
        geom_line(lwd = 1, color = '#00c9a1') +
        labs(title = 'Average steps of each interval',
         x = 'Interval',
         y = 'Steps') +
        facet_grid(weekd~.) +
        theme_bw()
```
